When Precision Therapy Creates Inprecise Choices

### 2022-12-15 Manuscript.qmd
revision

This question was the subject of recent reviews [@reinwald2018], and an expert guideline written by experts on behalf of the European Hematology Association (EHA) in cooperation with the Cochrane Hematology Group [@stemler_etal22]. For the resulting guidelines, studies including 19 relevant novel targeted agents currently approved or in later stages of clinical trials were reviewed, summarized and depending on available evidence, prepared as consensus-based recommendations despite the lack of high-quality studies assessing the use of antifungal agents in this setting [@stemler_etal22; @stemler_cornely22]. A summary of the recommendations for key approved agents is provided in [@tbl-drugs], the continued use of posaconazole prophylaxis was still recommended if patients are receiving targeted therapies as part of standard RIC. However, in consolidation chemotherapy, the risk of IFD is generally lower especially for invasive aspergillosis unless the patient has persistent neutropenia or prior fungal infections [@pagano_etal16], therefore continued use of posaconazole prophylaxis in patients received targeted therapies as maintenance or consolidation chemotherapy is conditional on the patient-specific risks. In this respect, we have found individualized risk stratification tools, such as institutionally-validated risk scores for aspergillosis [@stanzani_etal19;@stanzani_lewis18]to be helpful for decisions of whether continued prophylaxis is necessary.

The third key criteria supporting the use of prophylaxis is diagnostic difficulty. The availability of improved diagnostic tools (serum galactomannan, PCR) and high-resolution CT scanning has improved the early diagnosis of these infections allowing for possible or a preemeptive treatment approaches- i.e. initiation of antifungal therapy after detection of early fungal biomarkers and suggestive chest CT findings. Preemptive management approaches for IFD have been shown to be as effective as empirical treatment in high-risk patients with persistent fever [@morrissey2013]. An EORTC-lead randomized trial demonstrated that preemptive antifungal strategy was safe for high-risk neutropenic patients given fluconazole as prophylaxis, halving the number of patients receiving mould-active antifungals without excess mortality or IFDs . However, primary antifungal prophylaxis with posaconazole has not been directly compared with empiric or preemptive therapy in a randomized clinical trial during RIC for AML.

In allogeneic HSCT recipients, a preemptive-like approach (fluconazole prophylaxis plus galactomannan monitoring) was proven to be as safe and effective as voriconazole prophylaxis both in terms of 6-month fungal-free survival (FFS) and relapse-free and overall survival {Wingard et al., 2010, #305904}. However, in patients transplanted for AML, patients who received voriconazole prophylaxis versus the preemptive approach experienced significantly fewer cases of IFD (8.5% vs. 21%, P=0.04) and had improved FFS (78% vs. 61%, P=0.04) and a trend towards improved overall survival (81% vs. 72%, P=0.32) {Wingard et al., 2010, #305904}. Therefore, the equivalence of preemptive versus prophylactic approach for the highest groups remains unproven {Coussement et al., 2021, #228156}. However, the safety of a preemptive approach in lieu of continued prophylaxis seems feasible in patients on SMKIs during consolidation phase, provided there is rapid availability to CT scanning and mycological testing, including when patients are in the outpatient setting where the bulk of these treatments are now being administered.

Collectively, the evidence does not seem to currently support abandoning primary antifungal prophylaxis in AML patients receiving SMKIs during RIC when the risk of IFD is highest. However, primary antifungal prophylaxis mold-active triazoles may be unnecessary for patients receiving lower- intensity therapy with a hypomethylating agent (decitabine or azacitidine) or low-dose cytarabine alone combined with drugs such as venetoclax unless extenuating risk factors are present, such as prolonged neutropenia {Nucci and Anaissie, 2014, #67476}. In these patients a preemptive or empiric management approach may be appropriate in lieu of continued prophylaxis to reduce the risk of toxicity and drug interactions with the SMKIs.

@@

Non-triazole based agents, namely echinocandins or liposomal amphotericin B, are often considered as alternatives to triazoles in patients who require SMKIs or are enrolled in clinical trials with new targeted therapies that preclude use triazoles because of QT-prolongation risk {Brüggemann et al., 2022, #246315}. Currently, only micafungin is approved for prophylaxis of Candida infection in patients undergoing allogeneic hematopoietic stem cell transplantation or patients who are expected to have neutropenia (absolute neutrophil count \< 500 cells / µl) for 10 or more days. Current guidelines for the treatment of invasive aspergillosis do not recommend echinocandin monotherapy as a frontline treatment for invasive aspergillosis due to limited evidence of efficacy in patients with probable or proven IA {Herbrecht et al., 2010, #83826; Ullmann et al., 2012, #93212; Viscoli et al., 2009, #67612}. A recent open-label randomized trial comparing caspofungin to fluconazole prophylaxis in children and adolescents undergoing treatment for AML found significantly lower cumulative incidence of proven or probable aspergillosis in patient who received caspofungin versus fluconazole (0.5% vs. 3%, P=0.04) {Fisher et al., 2019, #304611}, although the rates of IA in the fluconazole arm were lower than reported in previous adult studies that compared posaconazole suspension to fluconazole/itraconazole (1% vs. 7%, P=0.04){Cornely et al., 2007, #18388}. Some {Gomes et al., 2014, #19244} but not all{Epstein et al., 2018, #24798;El Cheikh et al., 2019, #101153} single-center observational studies examining echinocandin prophylaxis during RIC for AML have reported higher rates of breakthrough IFD in patients receiving echinocandin prophylaxis versus triazoles. Therefore, it is still unclear whether echinocandin prophylaxis, despite the improved safety profile when administered with SMKIs, is as effective as posaconazole prophylaxis for preventing invasive mold disease.

>>>>

This question was the subject of recent reviews [@reinwald2018], and an expert guideline written by experts on behalf of the European Hematology Association (EHA) in cooperation with the Cochrane Hematology Group [@stemler_etal22]. For the resulting guidelines, studies including 19 relevant novel targeted agents currently approved or in later stages of clinical trials were reviewed, summarized and depending on available evidence, prepared as consensus-based recommendations despite the lack of high-quality studies assessing the use of antifungal agents in this setting [@stemler_etal22; @stemler_cornely22]. A summary of the recommendations for key approved agents is provided in [@tbl-drugs], the continued use of posaconazole prophylaxis was still recommended if patients are receiving targeted therapies as part of standard RIC. However, in consolidation chemotherapy, the risk of IFD is generally lower especially for invasive aspergillosis unless the patient has persistent neutropenia or prior fungal infections [@pagano_etal16], therefore continued use of posaconazole prophylaxis in patients received targeted therapies as maintenance or consolidation chemotherapy is conditional on the patient-specific risks. In this respect, we have found individualized risk stratification tools, such as institutionally-validated risk scores for aspergillosis [@stanzani_etal19; @stanzani_lewis18]to be helpful for decisions of whether continued prophylaxis is necessary.

The third key criteria supporting the use of prophylaxis is diagnostic difficulty. The availability of improved diagnostic tools (serum galactomannan, PCR) and high-resolution CT scanning has improved the early diagnosis of these infections allowing for possible or a preemeptive treatment approaches- i.e. initiation of antifungal therapy after detection of early fungal biomarkers and suggestive chest CT findings. Preemptive management approaches for IFD have been shown to be as effective as empirical treatment in high-risk patients with persistent fever [@morrissey2013]. An EORTC-lead randomized trial demonstrated that preemptive antifungal strategy was safe for high-risk neutropenic patients given fluconazole as prophylaxis, halving the number of patients receiving mould-active antifungals without excess mortality or IFDs [@Maertens2022]. However, primary antifungal prophylaxis with posaconazole has not been directly compared with empiric or preemptive therapy in a randomized clinical trial during RIC for AML.

In allogeneic HSCT recipients, a preemptive-like approach (fluconazole prophylaxis plus galactomannan monitoring) was proven to be as safe and effective as voriconazole prophylaxis both in terms of 6-month fungal-free survival (FFS) and relapse-free and overall survival [@Wingard2010a]. However, in patients transplanted for AML, patients who received voriconazole prophylaxis versus the preemptive approach experienced significantly fewer cases of IFD (8.5% vs. 21%, P=0.04) and had improved FFS (78% vs. 61%, P=0.04) and a trend towards improved overall survival (81% vs. 72%, P=0.32). Therefore, the equivalence of preemptive versus prophylactic approach for the highest groups remains unproven [@Coussement2021]. However, the safety of a preemptive approach in lieu of continued prophylaxis seems feasible in patients on SMKIs during consolidation phase, provided there is rapid availability to CT scanning and mycological testing, including when patients are in the outpatient setting where the bulk of these treatments are now being administered.

Collectively, the evidence does not seem to currently support abandoning primary antifungal prophylaxis in AML patients receiving SMKIs during RIC when the risk of IFD is highest. However, primary antifungal prophylaxis mold-active triazoles may be unnecessary for patients receiving lower- intensity therapy with a hypomethylating agent (decitabine or azacitidine) or low-dose cytarabine alone combined with drugs such as venetoclax unless extenuating risk factors are present, such as prolonged neutropenia [@nucci_anaissie14] In these patients a preemptive or empiric management approach may be appropriate in lieu of continued prophylaxis to reduce the risk of toxicity and drug interactions with the SMKIs.

@@

Non-triazole based agents, namely echinocandins or liposomal amphotericin B, are often considered as alternatives to triazoles in patients who require SMKIs or are enrolled in clinical trials with new targeted therapies that preclude use triazoles because of QT-prolongation risk [@bruggemann_etal22]. Currently, only micafungin is approved for prophylaxis of Candida infection in patients undergoing allogeneic hematopoietic stem cell transplantation or patients who are expected to have neutropenia (absolute neutrophil count \< 500 cells / µl) for 10 or more days. Current guidelines for the treatment of invasive aspergillosis do not recommend echinocandin monotherapy as a frontline treatment for invasive aspergillosis due to limited evidence of efficacy in patients with probable or proven IA [@ullmann2012a; @viscoli2009; @herbrecht2010a]. A recent open-label randomized trial comparing caspofungin to fluconazole prophylaxis in children and adolescents undergoing treatment for AML found significantly lower cumulative incidence of proven or probable aspergillosis in patient who received caspofungin versus fluconazole (0.5% vs. 3%, P=0.04) [@fisher_etal19a], although the rates of IA in the fluconazole arm were lower than reported in previous adult studies that compared posaconazole suspension to fluconazole/itraconazole (1% vs. 7%, P=0.04) [@cornely2007J]. Some [@gomes2014] but not all [@Epstein2018a]{Epstein et al., 2018, #24798;El Cheikh et al., 2019, #101153} single-center observational studies examining echinocandin prophylaxis during RIC for AML have reported higher rates of breakthrough IFD in patients receiving echinocandin prophylaxis versus triazoles. Therefore, it is still unclear whether echinocandin prophylaxis, despite the improved safety profile when administered with SMKIs, is as effective as posaconazole prophylaxis for preventing invasive mold disease.
